---
title: "My Portfolio for Computational Musicologi"
author: "Jan Hengeveld"
date: "2023-02-25"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true

---

```{r, setup}
library(tidyverse)
library(plotly)
library(spotifyr)
library(compmus)
```


### Computational-Musicology

A project in my third year of doing a Bachelors Study in Musicology at the University of Amsterdam with a specialization into Music Cognition. Project for analysing and comparing music playlists on Spotify, specifically searching for difference in 'calming/relaxing music' for humans and for dogs


![](Home>MyPortfolio/dog-jamming-out-to-music(1).jpg)


***

#### The playlists used in Spotify:

#### [Spotify playlist humans](https://open.spotify.com/playlist/3B0FtfxNiFOo82o8lmJcIp?si=76769a4a14014af1)

## Project for analysing and comparing music on Spotify, specifically searching for similarities and/or differences in 'calming/relaxing music' for humans and for dogs 
## Portfolio
 
### Introduction to analysis of music, aimed at relaxations

#### For this project I have analysed playlists on Spotify which are named either 'calming' or 'relaxing' or 'relaxation' and group them (eventually after some further research) fourfold: 
1. general playlist for humans, 
2. general playlist for dogs
3. specific playlist for humans, based upon research
4. specific playlist for dogs, based upon research

Which exact music elements are relevant for calming dogs may have been been applied. 

The corpus has been made from songs on many playlists that exist for those groups. It is my working hypothesis that the first two groups are hardly different and that a high level of antropomorphism is applicable => what humans define and perceive as relaxing will be true for dogs also. Group 3 eand 4 require further analysis 

Potentially a third group, well defined on scientific research, may show significant differences on a variety of elements. It might also be that 'true' calming music for dogs is based on elements that disqualify for humans as true music (i.e. pitches in sounds at frequency levels unhearable for humans but hearable for dogs. 

Re. the 3 group: see https://icalmpet.com/wp-content/uploads/BioAcoustic-Research-and-Development-Executive-Summary.pdf

I based my third group , called "Through a Dogs Ear, science-based supposedly. upon this paper and the available public playlists in Spotify labeled "Through a Dogs Ear" (published by authors of this research).

** Research has been done into calming/relaxing effects of music, both for humans as for dogs. For the first group, humans, we know much more given a higher quality level of feedback. For dogs, research shows that similar aspects apply as for humans (tempo, loudness, pitch, instruments) but also differences (variety, genre, 'nature'sounds).
At this stage I haven't identified nor selected specific tracks for each of these 2 or 3 groups. I need to do more analysis of previous research into calming music for dogs to identify the musical elements that appear to be relevant.

test
### A global analysis of 4 relaxation playlists: humans, dogs, 'sciencehumans' and sciencedogs'
#### bla bla bla

```{r, setup}

 test5 |>
    ggplot(aes(x = energy)) +
    geom_histogram(binwidth = 0.1) +
    facet_wrap(~category)

```



### Further analysise

#### I selected the song "Totally beached" from the third group (Dogs, science-based) for a further analysis because it has a strong Regae 'vibe' to it and other research on the effect of music on calming dogs found that Regae music was preferred by dogs (reference to article) 

#### I selected the song "Weightless" by Marconi Union based upon the following research:  https://www.britishacademyofsoundtherapy.com/wp-content/uploads/2019/10/Mindlab-Report-Weightless-Radox-Spa.pdf. This study 

#### either 'calming' or 'relaxing' or 'relaxation' and group them twofold or threefold: 1 group general and 1 group aimed at dogs. Possibly also a third group for dogs where scientific research (into which exactc music elements are relevant for calming dogs) may have been been applied. The corpus can be made from songs on many playlists that exist for those groups. It is my working hypothesis that the first two groups are hardly different and that a high level of antropomorphism is applicable => what humans define and perceive as relaxing will be true for dogs also. Potentially a third group, well defined on scientific research, may show significant differences on a variety of elements. It might also be that 'true' calming music for dogs is based on elements that disqualify for humans as true music (i.e. pitches in sounds at frequency levels unhearable for humans but hearable for dogs. 
** Research has been done into calming/relaxing effects of music, both for humans as for dogs. For the first group, humans, we know much more given a higher quality level of feedback. For dogs, research shows that similar aspects apply as for humans (tempo, loudness, pitch, instruments) but also differences (variety, genre, 'nature'sounds).
At this stage I haven't identified nor selected specific tracks for each of these 2 or 3 groups. I need to do more analysis of previous research into calming music for dogs to identify the musical elements that appear to be relevant.
